SSwAB-IT Phase 1 Analysis

### 2024-09-19 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

@@

For these agreement analyses, we reported accuracy metrics that include all samples. Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative). Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement [@mchughInterraterReliabilityKappa2012;@ranganathanCommonPitfallsStatistical2017]. A Cohen's kappa of NaN or NA indicates a lack of variability between the two tests.

@@

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  colnames(d) <- c("COS", "SOS")
  d <- d[complete.cases(d), ]
  # Agreement metrics
  c <- confusionMatrix(d$SOS,
    reference = d$COS,
    positive = "Positive", mode = "everything",
    dnn = c("SOS", "COS")
  )
  # Print
  cat("\n")
  cat("####", "2x2 Table")
  cat("\n")
  d %>%
    tbl_cross(percent = "cell") %>%
    print()
  cat("\n")
  cat("####", "2x2 Table (With CIs)")
  cat("\n")
  d %>%
    tbl_cross(percent = "cell") %>%
    add_ci() %>%
    print()
  cat("\n")
  cat("Overall accuracy (both negative/negative and positive/positive) was ",
    scales::percent(c$overall["Accuracy"], accuracy = 0.1), " (95% CI: ",
    scales::percent(c$overall["AccuracyLower"], accuracy = 0.1), " - ",
    scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1), ").",
    sep = ""
  )
  cat("\n")
  cat("\n")
  cat("Kappa was ", round(c$overall["Kappa"], 3), ".", sep = "")
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of negative results")
  cat("\n")
  print(plot(euler(d == "Negative"), quantities = list(cex = 3)))
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of positive results")
  cat("\n")
  print(plot(euler(d == "Positive"), quantities = list(cex = 3)))
  cat("\n")
}
agreement_test <- function(bug, group = "group", data = df) {
  # Prepare data
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos, group)]
  d <- d[complete.cases(d), ]
  colnames(d) <- c("COS", "SOS", "Group")
  # Agreement and detailed agreement
  d$agree <- d$COS == d$SOS
  d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
  # Print
  cat("\n")
  d %>%
    select(agree, detailed_agree, Group) %>%
    tbl_summary(by = Group, label = list(
      agree ~ "Agree (Both Positive/Positive and Negative/Negative)",
      detailed_agree ~ "Detailed Agreement"
    )) %>%
    add_overall() %>%
    add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e6)) %>%
    separate_p_footnotes() %>%
    print()
  cat("\n")
}
```

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("##", as.character(bug_list[b]))
  cat("\n")
  cat("###", "Overall agreement")
  cat("\n")
  agreement(b)
  cat("\n")
  cat("###", "Agreement by group")
  cat("\n")
  agreement_test(b)
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by age group")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by gender assigned at birth")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by Azithromycin status")
  cat("\n")
  agreement_test(b, group = "azithromycin")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Tobramycin status")
  cat("\n")
  agreement_test(b, group = "inhaled_tobramycin")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Aztreonam status")
  cat("\n")
  agreement_test(b, group = "inhaled_aztreonam")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Colisitin status")
  cat("\n")
  agreement_test(b, group = "inhaled_colisitin")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Vancomycin status")
  cat("\n")
  agreement_test(b, group = "inhaled_vancomycin")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by other inhaled antibiotics status")
  cat("\n")
  agreement_test(b, group = "other_inhaled_antibiotics")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by chronic oral antibiotic status")
  cat("\n")
  agreement_test(b, group = "chronic_oral_antibiotic")
  cat("\n")
}))
```

@@

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by CFTR modulator status")
  cat("\n")
  agreement_test(b, group = "cftr_modulator")
  cat("\n")
}))
```

# References

>>>>

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

@@

For these agreement analyses, we reported accuracy metrics that include all samples. Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative). Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analogously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement. A Cohen's kappa of NaN or NA indicates a lack of variability between the two tests.

@@

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  colnames(d) <- c("COS", "SOS")
  d <- d[complete.cases(d), ]
  # Agreement metrics
  c <- confusionMatrix(d$SOS,
    reference = d$COS,
    positive = "Positive", mode = "everything",
    dnn = c("SOS", "COS")
  )
  # Print
  cat("\n")
  cat("####", "2x2 Table")
  cat("\n")
  d %>%
    tbl_cross(percent = "cell") %>%
    print()
  cat("\n")
  cat("####", "2x2 Table (With CIs)")
  cat("\n")
  d %>%
    tbl_summary(by = COS, percent = "cell") %>%
    add_ci(method = list(all_categorical() ~ "wald")) %>%
    print()
  cat("\n")
  cat("Overall accuracy (both negative/negative and positive/positive) was ",
    scales::percent(c$overall["Accuracy"], accuracy = 0.1), " (95% CI: ",
    scales::percent(c$overall["AccuracyLower"], accuracy = 0.1), " - ",
    scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1), ").",
    sep = ""
  )
  cat("\n")
  cat("\n")
  cat("Kappa was ", round(c$overall["Kappa"], 3), ".", sep = "")
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of negative results")
  cat("\n")
  print(plot(euler(d == "Negative"),
    quantities = list(cex = 3),
    fills = list(fill = c("#FFC107", "#1E88E5"))
  ))
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of positive results")
  cat("\n")
  print(plot(euler(d == "Positive"),
    quantities = list(cex = 3),
    fills = list(fill = c("#FFC107", "#1E88E5"))
  ))
  cat("\n")
}
agreement_test <- function(bug, group = "group", data = df) {
  # Prepare data
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos, group)]
  d <- d[complete.cases(d), ]
  colnames(d) <- c("COS", "SOS", "Group")
  # Agreement and detailed agreement
  d$agree <- d$COS == d$SOS
  d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
  # Print
  cat("\n")
  d %>%
    select(agree, detailed_agree, Group) %>%
    tbl_summary(by = Group, label = list(
      agree ~ "Agree (Both Positive/Positive and Negative/Negative)",
      detailed_agree ~ "Detailed Agreement"
    )) %>%
    add_overall() %>%
    add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e6)) %>%
    separate_p_footnotes() %>%
    print()
  cat("\n")
}
```

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  print(b)
  cat("\n")
  cat("##", as.character(bug_list[b]))
  cat("\n")
  cat("###", "Overall agreement")
  cat("\n")
  agreement(b)
  cat("\n")
  cat("###", "Agreement by group")
  cat("\n")
  agreement_test(b)
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by age group")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by gender assigned at birth")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by Azithromycin status")
  cat("\n")
  agreement_test(b, group = "azithromycin")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Tobramycin status")
  cat("\n")
  agreement_test(b, group = "inhaled_tobramycin")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Aztreonam status")
  cat("\n")
  agreement_test(b, group = "inhaled_aztreonam")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Colisitin status")
  cat("\n")
  agreement_test(b, group = "inhaled_colisitin")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Vancomycin status")
  cat("\n")
  agreement_test(b, group = "inhaled_vancomycin")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by other inhaled antibiotics status")
  cat("\n")
  agreement_test(b, group = "other_inhaled_antibiotics")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by chronic oral antibiotic status")
  cat("\n")
  agreement_test(b, group = "chronic_oral_antibiotic")
  cat("\n")
}))
```

@@

```{r results='asis'}
#| message: false
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by CFTR modulator status")
  cat("\n")
  agreement_test(b, group = "cftr_modulator")
  cat("\n")
}))
```
